Parallelization of Fast $\ell_1$-Minimization for Face Recognition

### 2011-03-22 main.tex
added caption
revised the experimental section

\subsection{CPU Implementation} Our CPU implementation of Algorithm
\ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
nicely onto the BLAS API (such as SGEMV, SAXPY {\bf? Victor, fill in here}), we
call Intel's MKL BLAS library, which internally takes advantage of the SSE
units on all available CPU cores. 
{\bf TODO: Our current implementation links against the MKL
library provided internally with matlab; we are in the process of updating the code to link directly against a newer version
of MKL}

@@

Operations that do not map well onto the BLAS API (mostly combined
vector-vector operations) were implemented as gpu kernels (in C for CUDA).
Even for operations that could have been implemented via multiple BLAS calls,
this was found to have better performance due to reducede bandwidth and kernel
call overhead {\bf Victor, verify or correct this}.

@@

The first experiment compares the performance of our two proposed implementations of 
the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
vector $\xx_0$.  
The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
$10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
Because the ground truth signal $\xx_0$ is known, the algorithm terminates
when $norm(\bx-\bx_0) < tolerance$.
The measurement vector was generated with the equation $\y_0 = A \x_0$.   
%We also varied the different inner and outer tolerance levels from 1e-2 to 1e-5.
We averaged the time taken for the ALM algorithm to solve 10 instances of a problem.

The results of this benchmark can be found in Figure \ref{fig:random_data}.
The x-axis represents the size of the $A$ matrix and the y-axis represents the
average amount of time it to complete a single problem instance.  A can be seen
in Figure \ref{fig:random_data}, the GPU implementation tends to be faster than
the CPU implementation at solving a single large problem, whereas the the CPU
implementation is faster at solving a single small problem.  
\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{results/random_data/time_vs_matrix_size_constant_tol}
\end{center}
\caption{Speed vs Dimensions of A.  Outer loop tolerance = 0.1.  Inner loop tolerance = $10^{-5}$}
\label{fig:random_data}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_2}
\end{center}
\caption{\bf CAPTION NEEDED}
\label{fig:random_data2}
\end{figure}

>>>>

\subsection{CPU Implementation} Our CPU implementation of Algorithm
\ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
nicely onto the BLAS API (such as SGEMV, SAXPY), we
call Intel's MKL BLAS library, which internally takes advantage of the SSE
units on all available CPU cores. 
{\bf TODO: Our current implementation links against the MKL
library provided internally with matlab; we are in the process of updating the code to link directly against a newer version
of MKL}

@@

Operations that do not map well onto the BLAS API (mostly combined
vector-vector operations) were implemented as gpu kernels (in C for CUDA).
Even for operations that could have been implemented via multiple BLAS calls,
this was found to have better performance due to reduced bandwidth and kernel
call overhead {\bf Victor, verify or correct this.  I'm not sure, currently asking Mark}.

@@

The first experiment compares the performance of our two proposed implementations of 
the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
vector $\xx_0$.  
The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
$10:1$ to $10:5$ with $d$ varying from 500 to 8000.
The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
Because the ground truth signal $\xx_0$ was known, the algorithm terminated
when $norm(\bx-\bx_0) < tolerance$.
The measurement vector was generated with the equation $\y_0 = A \x_0$.

The results of this benchmark can be found in Figure \ref{fig:random_data}.
The x-axis represents the size of the $A$ matrix and the y-axis represents the
average amount of time it to complete a single problem instance.  The GPU implementation tends to be faster than
the CPU implementation at solving a single large problem, whereas the the CPU
implementation is faster at solving a single small problem.  Figure \ref{fig:random_data2} shows the various sizes and tolerances at which our GPU implementation
begins to run faster than the CPU implementation on synthetic data and will provide researchers 
with a starting point of where to begin setting their tolerance levels for the ALM algorithm.
\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{results/random_data/time_vs_matrix_size_constant_tol}
\end{center}
\caption{Speed vs Dimensions of A.  Outer loop tolerance = 0.1.  Inner loop tolerance = $10^{-5}$}
\label{fig:random_data}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_3}
\end{center}
\caption{Relation between CPU and GPU runtime vs tolerance level}
\label{fig:random_data2}
\end{figure}
